Understanding and Verifying the Shape of Stencil Computations

### 2017-03-22 oopsla17/lattice-model.tex
what I got

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage{indentfirst}
\usepackage{cleveref}

@@

\newcommand{\zinf}{\textnormal{$\mathbb{Z}^\infty$}}
\newcommand{\interp}[1]{\llbracket\texttt{#1}\rrbracket}
\newcommand{\interv}[3]{\textnormal{$\lbrack#1,#2\rbrack^{#3}$}}
\newcommand{\vecgen}[3]{\textnormal{$\textit{vec-gen}_{#1}(#2,#3)$}}

@@

\begin{defn}{(\emph{Offset space})}
  \emph{Offset space} is a subset of \zinf{} members. $\mathcal{P}(\zinf{})$
  couple with $\supseteq$ is the power set distributive lattice.
\end{defn}

\begin{defn}{(\emph{Offsets vector})}
  An \emph{offsets vector} is a $N$ dimensional vector of offset spaces denoted
  by $(\textit{offs}_1, \dots, \textit{offs}_N)$.

  The set of $N$ dimensional offsets vectors is $\mathcal{P}{(\zinf{})}^N$. We
  define the partial order $\leq$ on $N$ dimensional offsets vectors
  \textit{voffs} and \textit{voffs'} as follows:
%
  \begin{equation*}
    \textit{voffs} \leq \textit{voffs'} \triangleq
      \forall n. 1
        \leq i \leq N \implies \textit{voffs}_i \supseteq \textit{voffs'}_i
  \end{equation*}
\end{defn}

\begin{prop}{}
  $(\mathcal{P}{(\zinf{})}^N, \leq)$ is a distributive lattice.
\end{prop}
%
\begin{proof}
  Product of two distributive lattices is itself a distributive lattice when the
  partial order is defined as componentwise comparison in the underlying
  lattice.
\end{proof}

\begin{defn}{(\emph{\zinf{} interval})}
  A \emph{\zinf{} interval} is of the form \interv{a}{b}{c}, where $a$ and $b$
  are drawn from \zinf{} with $a \leq 0 \leq b$ and $c$ is drawn from $\{
  \textsf{p}, \textsf{np} \}$. We define it as follows:
%
  \begin{equation*}
    \interv{a}{b}{c} \triangleq
      \{ n \;|\; a \leq n \leq b \wedge (c = \textsf{np} \implies n \neq 0) \}
  \end{equation*}

@@

\begin{prop}{}
  $(\textit{Interval}, \supseteq)$ is a sublattice of
  $(\mathcal{P}(\zinf{}),\supseteq)$ with \interv{-\infty}{\infty}{\textsf{p}}
  and \interv{0}{0}{\textsf{np}} as its top and bottom elements.
\end{prop}
%
\begin{proof}
  We know $\textit{Interval} \subseteq \mathcal{P}(\zinf{})$ since intervals
  represents sets drawn from {\zinf{}}

  Let \interv{a}{b}{c} and \interv{d}{e}{f} be two arbitrary \zinf{} intervals.
  We need to show that their meet and join both lie in \textit{Interval}.
%
  \begin{align*}
    \interv{a}{b}{c} \wedge \interv{d}{e}{f} = &
      \{ n \;|\; a \leq n \leq b \wedge (c = \textsf{np} \implies n \neq 0) \}
      \;\cup \\
      & \{ n \;|\; d \leq n \leq e \wedge (f = \textsf{np} \implies n \neq 0) \}
      \\
    = & \{ n \;|\; \min{a,d} \leq n \leq \max{b,e} \wedge ((c =
      \textsf{np} \wedge f = \textsf{np}) \implies n \neq 0) \} \\
    = & \interv{\min{\{a,d\}}}{\max{\{b,e\}}}{k}
  \end{align*}
%
  where $k$ is \textsf{np} if both $c$ and $d$ are \textsf{np} and \textsf{p}
  otherwise. Minimum of lower bound is still negative or zero and maximum of
  upper bounds is still positive or zero, hence the meet lies in
  \textit{Interval}.
%
  \begin{align*}
    \interv{a}{b}{c} \vee \interv{d}{e}{f} = &
      \{ n \;|\; a \leq n \leq b \wedge (c = \textsf{np} \implies n \neq 0) \}
      \;\cap \\
      & \{ n \;|\; d \leq n \leq e \wedge (f = \textsf{np} \implies n \neq 0) \}
      \\
    = & \{ n \;|\; \max{a,d} \leq n \leq \min{b,e} \wedge ((c =
      \textsf{np} \vee f = \textsf{np}) \implies n \neq 0) \} \\
    = & \interv{\max{\{a,d\}}}{\min{\{b,e\}}}{k}
  \end{align*}
%
  where $k$ is \textsf{np} if either $c$ or $d$ are \textsf{np} and \textsf{p}
  otherwise. Maximum of lower bound is still negative or zero and minimum of
  upper bounds is still positive or zero, hence the join lies in
  \textit{Interval}.

  \interv{-\infty}{\infty}{\textsf{p}} is then \zinf{} and
  \interv{0}{0}{\textsf{np}} is $\emptyset$. Since these are in the sublattice
  and they are the bottom and top elements in the parent lattice, they are
  necessarily the bottom and top elements in the sublattice.
\end{proof}

\begin{prop}{}
  $(\textit{Interval},\supseteq)$ is a distributive lattice.
\end{prop}
%
\begin{proof}
  It is enough to show for arbitrary $\interv{a}{b}{c}$, $\interv{d}{e}{f}$,
  $\interv{g}{h}{i}$ in \textit{Interval}, $\interv{a}{b}{c} \vee
  (\interv{d}{e}{f} \wedge \interv{g}{h}{i}) = (\interv{a}{b}{c} \vee
  \interv{d}{e}{f}) \wedge (\interv{a}{b}{c} \vee \interv{g}{h}{i})$. When this
  holds the dual law also holds. The property immediately follows from
  definition due to distributivity of set intersection over union.
\end{proof}
%
\begin{defn}{(\emph{Vector interval})}
  A vector interval is an $N$ dimensional object which has \zinf{} intervals at
  each dimension. It is denoted with $(\textit{int}_1, \dots, \textit{int}_n)$,
  where $\textit{int}_i$ is a \zinf{} interval. If \textit{vint} is a vector
  interval, $\textit{vint}_i$ denotes the \zinf{} interval at the $i^{th}$
  dimension.

  The set of $N$ dimensional vector intervals is $\textit{Interval}^N$. We
  define a partial order $\leq$ on $N$ dimensional vector intervals
  \textit{vint} and \textit{vint'} as follows:
%
  \begin{equation*}
    \textit{vint} \leq \textit{vint}' \triangleq
      \forall n. 1
        \leq i \leq N \implies \textit{vint}_i \supseteq \textit{vint'}_i
  \end{equation*}
\end{defn}
%
\begin{prop}{}
  $(\textit{Interval}^N, \leq)$ is a distributive lattice.
\end{prop}
%
\begin{proof}
  It is a standard result that product of two distributive lattices is a
  distributive lattice itself when the partial order on the resulting lattice is
  defined as the componentwise comparison. This generalises to $N$ dimensions by
  induction.
\end{proof}

@@

We give some basic interpretations: $\interp{\texttt{n}} = n$ where $n$ is in
$\mathbb{Z}$, $\interp{pointed} = \llbracket \epsilon \rrbracket = \textsf{p}$,
$\interp{\texttt{nonpointed}} = \textsf{np}$. Typographically, cursive font is
used within specifications to indicate a partial textual capture.

@@

First, we give meaning to region constants (introduced in \Cref{}).
%
\begin{align*}
  \interp{\texttt{pointed(dim=\textit{i})}}_N & =
    \vecgen{N}{\interp{\textit{i}}}{\interv{0}{0}{\textsf{p}}}\\
%
  \interp{\texttt{centered(dim=\textit{i}, depth=\textit{k}, \textit{p})}}_N & =
    \vecgen{N}
           {\interp{\textit{i}}}
           {\interv{-\interp{\textit{k}}}
                   {\interp{\textit{k}}}
                   {\interp{\textit{p}}}
           } \\
%
  \interp{\texttt{forward(dim=\textit{i}, depth=\textit{k}, \textit{p})}}_N & =
    \vecgen{N}
           {\interp{\textit{i}}}
           {\interv{0}
                   {\interp{\textit{k}}}
                   {\interp{\textit{p}}}
           } \\
%
  \interp{\texttt{backward(dim=\textit{i}, depth=\textit{k}, \textit{p})}}_N & =
    \vecgen{N}
           {\interp{\textit{i}}}
           {\interv{-\interp{\textit{k}}}
                   {0}
                   {\interp{\textit{p}}}
           }
\end{align*}

@@

\section{Equational theory}

>>>>

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage{indentfirst}
\usepackage{cleveref}
\usepackage{fixltx2e}

@@

\newcommand{\zinf}{\textnormal{$\mathbb{Z}_\infty$}}
\newcommand{\interp}[1]{\llbracket\texttt{#1}\rrbracket}
\newcommand{\interv}[3]{\textnormal{$\lbrack#1,#2\rbrack^{#3}$}}
\newcommand{\vecgen}[3]{\textnormal{$\textit{vec-gen}_{#1}(#2,#3)$}}
\newcommand{\textcap}[1]{\texttt{\textit{#1}}}

@@

\begin{defn}{(\emph{Offset space})}
  \emph{Offset space} is a subset of \zinf{} members. It is used to capture
  offsets from induction variables.

%  $\mathcal{P}(\zinf{})$ couple with $\supseteq$ is the power set distributive
%  lattice.
\end{defn}

\begin{defn}{(\emph{\zinf{} interval})}
  A \emph{\zinf{} interval} is of the form \interv{a}{b}{c}, where $a$ and $b$
  are drawn from \zinf{} with $a \leq 0 \leq b$ and $c$ is a Boolean value. We
  define it as follows:
%
  \begin{equation*}
    \interv{a}{b}{c} \triangleq
      \{ n \;|\; a \leq n \leq b \wedge (\neg c \implies n \neq 0) \}
  \end{equation*}

@@

\begin{prop}{}
  We have the following identity for \zinf{} intervals:
%
  \begin{equation*}
    \interv{a}{b}{c} \cap \interv{d}{e}{f} =
      \interv{\max \{a,d\}}{\min \{b,e\}}{\neg c \vee \neg f}
  \end{equation*}\label{def:zinf-intersect}
\end{prop}
%
\begin{proof}
  \begin{align*}
    \interv{a}{b}{c} \cap \interv{d}{e}{f} = &
      \; \{ n \;|\; a \leq n \leq b \wedge (\neg c \implies n \neq 0) \}
      \;\cap \\
      & \; \{ n \;|\; d \leq n \leq e \wedge (\neg f \implies n \neq 0) \}
      \\
    = & \; \{ n \;|\; \max \{a,d\} \leq n \leq \min \{b,e\} \wedge (\neg c
      \vee \neg f \implies n \neq 0) \} \\
    = & \; \interv{\max{\{a,d\}}}{\min{\{b,e\}}}{\neg c \vee \neg f}
  \end{align*}
\end{proof}

%\begin{prop}{}
%  $(\textit{Interval}, \supseteq)$ is a sublattice of
%  $(\mathcal{P}(\zinf{}),\supseteq)$ with \interv{-\infty}{\infty}{\textsf{p}}
%  and \interv{0}{0}{\textsf{np}} as its top and bottom elements.
%\end{prop}
%%
%\begin{proof}
%  We know $\textit{Interval} \subseteq \mathcal{P}(\zinf{})$ since intervals
%  represents sets drawn from {\zinf{}}
%
%  Let \interv{a}{b}{c} and \interv{d}{e}{f} be two arbitrary \zinf{} intervals.
%  We need to show that their meet and join both lie in \textit{Interval}.
%%
%  \begin{align*}
%    \interv{a}{b}{c} \wedge \interv{d}{e}{f} = &
%      \{ n \;|\; a \leq n \leq b \wedge (c = \textsf{np} \implies n \neq 0) \}
%      \;\cup \\
%      & \{ n \;|\; d \leq n \leq e \wedge (f = \textsf{np} \implies n \neq 0) \}
%      \\
%    = & \{ n \;|\; \min{a,d} \leq n \leq \max{b,e} \wedge ((c =
%      \textsf{np} \wedge f = \textsf{np}) \implies n \neq 0) \} \\
%    = & \interv{\min{\{a,d\}}}{\max{\{b,e\}}}{k}
%  \end{align*}
%%
%  where $k$ is \textsf{np} if both $c$ and $d$ are \textsf{np} and \textsf{p}
%  otherwise. Minimum of lower bound is still negative or zero and maximum of
%  upper bounds is still positive or zero, hence the meet lies in
%  \textit{Interval}.
%%
%  \begin{align*}
%    \interv{a}{b}{c} \vee \interv{d}{e}{f} = &
%      \{ n \;|\; a \leq n \leq b \wedge (c = \textsf{np} \implies n \neq 0) \}
%      \;\cap \\
%      & \{ n \;|\; d \leq n \leq e \wedge (f = \textsf{np} \implies n \neq 0) \}
%      \\
%    = & \{ n \;|\; \max{a,d} \leq n \leq \min{b,e} \wedge ((c =
%      \textsf{np} \vee f = \textsf{np}) \implies n \neq 0) \} \\
%    = & \interv{\max{\{a,d\}}}{\min{\{b,e\}}}{k}
%  \end{align*}
%%
%  where $k$ is \textsf{np} if either $c$ or $d$ are \textsf{np} and \textsf{p}
%  otherwise. Maximum of lower bound is still negative or zero and minimum of
%  upper bounds is still positive or zero, hence the join lies in
%  \textit{Interval}.
%
%  \interv{-\infty}{\infty}{\textsf{p}} is then \zinf{} and
%  \interv{0}{0}{\textsf{np}} is $\emptyset$. Since these are in the sublattice
%  and they are the bottom and top elements in the parent lattice, they are
%  necessarily the bottom and top elements in the sublattice.
%\end{proof}

%\begin{prop}{}
%  $(\textit{Interval},\supseteq)$ is a distributive lattice.
%\end{prop}
%%
%\begin{proof}
%  It is enough to show for arbitrary $\interv{a}{b}{c}$, $\interv{d}{e}{f}$,
%  $\interv{g}{h}{i}$ in \textit{Interval}, $\interv{a}{b}{c} \vee
%  (\interv{d}{e}{f} \wedge \interv{g}{h}{i}) = (\interv{a}{b}{c} \vee
%  \interv{d}{e}{f}) \wedge (\interv{a}{b}{c} \vee \interv{g}{h}{i})$. When this
%  holds the dual law also holds. The property immediately follows from
%  definition due to distributivity of set intersection over union.
%\end{proof}

\begin{defn}{(\emph{Offsets vector})}
  An \emph{offsets vector} is a $N$ dimensional vector of offset spaces denoted
  by $(\textit{offs}_1, \dots, \textit{offs}_N)$ to mean $\textit{offs}_1 \times
  \dots \times \textit{offs}_N$.

  The set of $N$ dimensional offsets vectors is
  $\mathcal{P}{(\zinf{}^{\mkern-19mu{N}})}$ and they trivially form a
  distributive lattice under the superset relation.
\end{defn}

\begin{defn}{(\emph{Vector interval})}
  A vector interval is an $N$ dimensional object which has \zinf{} intervals at
  each dimension. It is denoted with $(\textit{int}_1, \dots, \textit{int}_n)$,
  where $\textit{int}_i$ is a \zinf{} interval. If \textit{vint} is a vector
  interval, $\textit{vint}_i$ denotes the \zinf{} interval at the $i^{th}$
  dimension.

  The set of $N$ dimensional vector intervals is $\textit{Interval}^N$.
\end{defn}

\begin{lem}{}
  Intersection distributes over vector intervals.
%
  \begin{equation*}
    (\textit{vint}_1, \dots, \textit{vint}_N) \;\cap\;
      (\textit{vint'}_1, \dots, \textit{vint'}_N)
      =
    (\textit{vint}_1 \;\cap\; \textit{vint'}_1, \dots,
     \textit{vint}_N \;\cap\; \textit{vint'}_N)
  \end{equation*}\label{lem:vector-dist}
\end{lem}
%
\begin{prop}{}
  We have the following identity for $N$ dimensional vector intervals:
%
  \begin{gather*}
    (\interv{a_1}{b_1}{c_1}, \dots, \interv{a_N}{b_N}{c_N}) \; \cap \;
      (\interv{a'_1}{b'_1}{c'_1}, \dots, \interv{a'_N}{b'_N}{c'_N}) \\
    = \\
    (\interv{\max \{a_1,a'_1\}}{\min \{b_1,b'_1\}}{\neg c_1 \vee \neg c'_1},
      \dots, \interv{\max \{a_N,a'_N\}}{\min \{b_N,b'_N\}}{\neg c_N \vee \neg c'_N})
  \end{gather*}
\end{prop}
%
\begin{proof}
  This immediately follows from \Cref{lem:vector-dist} and
  \Cref{def:zinf-intersect}.
\end{proof}

\begin{defn}{(\emph{Region space})}
  Region space defines the set \textit{Region} in $N$ dimensions inductively as
  follows:
%
  \begin{enumerate}
    \item If $R$ is in $\textit{Interval}^N$, then $R$ is in \textit{Region}.
    \item If $R$ and $S$ are in \textit{Region}, then so are $R \cap S$ and $R
      \cup S$.
  \end{enumerate}
\end{defn}

\begin{prop}{}
  $(\textit{Region},\supseteq)$ is a distributive lattice.
\end{prop}
%
\begin{proof}
  Any sublattice of a distributive lattice is a distributive lattice. Power sets
  coupled with partially ordered with superset or equals relation are
  distributive lattices. \textit{Region} is a subset of
  $\mathcal{P}{(\zinf{}^{\mkern-19mu{N}})}$ since all vector intervals are
  subsets of this power set and union and intersection cannot add an element
  that is not already in the powerset. Therefore, \textit{Region} is a
  distributive sublattice of $\mathcal{P}{(\zinf{}^{\mkern-19mu{N}})}$.
\end{proof}

@@

We give some basic interpretations: $\interp{\texttt{n}} = n$ where $n$ is in
$\mathbb{Z}$, $\interp{pointed} = \llbracket \epsilon \rrbracket =
\textsf{true}$, $\interp{\texttt{nonpointed}} = \textsf{false}$.
Typographically, cursive font is used within specifications to indicate a
partial textual capture.

@@

First, we give meaning to region constants (introduced in \Cref{}).
%
\begin{align*}
  \interp{\texttt{pointed(dim=\textit{i})}}_N & =
    \vecgen{N}{\interp{\textit{i}}}{\interv{0}{0}{\textsf{true}}}\\
%
  \interp{\texttt{centered(dim=\textit{i}, depth=\textit{k}, \textit{p})}}_N & =
    \vecgen{N}
           {\interp{\textit{i}}}
           {\interv{-\interp{\textit{k}}}
                   {\interp{\textit{k}}}
                   {\interp{\textit{p}}}
           } \\
%
  \interp{\texttt{forward(dim=\textit{i}, depth=\textit{k}, \textit{p})}}_N & =
    \vecgen{N}
           {\interp{\textit{i}}}
           {\interv{0}
                   {\interp{\textit{k}}}
                   {\interp{\textit{p}}}
           } \\
%
  \interp{\texttt{backward(dim=\textit{i}, depth=\textit{k}, \textit{p})}}_N & =
    \vecgen{N}
           {\interp{\textit{i}}}
           {\interv{-\interp{\textit{k}}}
                   {0}
                   {\interp{\textit{p}}}
           }
\end{align*}

@@

\section{Equational \& approximation theories}

We know introduce the equational and approximation theory of stencil
specifications based on $\equiv$ and $<\:$.

\subsection{Equivalences}

\begin{description}
  \item[Basic] \texttt{*} and \texttt{+} are both idempotent, commutative, and
    associative;
%
  \item[Dist] \texttt{*} distributed over \texttt{+} and dually
    \texttt{+} distributes over \texttt{*}, meaning if \textcap{R}, \textcap{S},
    and \textcap{T} are regions, then we have the following equivalences:
%
    \begin{align*}
      \texttt{\textcap{R}*(\textcap{S}+\textcap{T})} & \equiv
        \texttt{(\textcap{R}*\textcap{S})+(\textcap{R}*\textcap{T})} &
      \texttt{\textcap{R}+(\textcap{S}*\textcap{T})} & \equiv
        \texttt{(\textcap{R}+\textcap{S})*(\textcap{R}+\textcap{T})}
    \end{align*}
%
  \item[Overlapping pointed] If \textcap{R} is one of \texttt{forward},
    \texttt{backward}, or \texttt{centered}, then we have
    $\texttt{\textcap{R}(depth=\textcap{k},dim=\textcap{n},\_) +
    pointed(dim=\textcap{n})} \equiv
    \texttt{\textcap{R}(depth=\textcap{k},dim=\textcap{n})}$.
%
  \item[Centered] The region constants \texttt{forward} and \texttt{backward}
    are two halves of \texttt{centered} specifications:
%
    \begin{gather*}
      \texttt{centered(depth=\textcap{k},dim=\textit{n},\textcap{p1})} \\
      \equiv \\
      \texttt{forward(depth=\textcap{k},dim=\textit{n},\textcap{p2}) +
        backward(depth=\textcap{k},dim=\textit{n},\textcap{p3})}
    \end{gather*}
%
    Here \textcap{p1} is \texttt{nonpointed} if both \textcap{p1} and
    \textcap{p3} are \texttt{nonpointed} and \texttt{pointed} otherwise.
\end{description}

\subsection{Approximations}
